Psychophysics of musical elements in the discrete-time representation of sound

### 2013-09-19 article/article.tex
feita intro do artigo

This paper purposes to represent musical structures and artifices by their discrete sound characteristics. The results include mathematical relations and their computer program implementations. Despite the general interest and number of knowledge areas involved, there are only a few number of books and computer implementations that tackle the subjects presented in this work. However, they mainly focus on aspects of musical sounds and ways to mimic traditional instruments. A detailed compilation of them is pointed out in the bibliography~\cite{dissertacao}. To the best of the author's knowledge, there is a lack of articles on the topic. Moreover, although current computer implementations use the analytical descriptions presented in this study in a implicitly manner, it seems that there is not a concise and mathematical description of the processes implemented, as they aim here to be further used as libraries for sound and music.

In order to address the concise description of musical elements and structures, the objectives of this paper are:

@@

\item Introduce a framework of sound synthesis with control at sample level and with potential uses in psycho-acoustical experiments and high-fidelity synthesis. The didactic presentation of the content favors its apprehension and usage. It is worthwhile to mention that this subject comprises diverse topics on signal processing, music and psycho-acoustics.

\item Provide the accessibility of the developed framework. The topics presented in this article are implemented as scripts, i.e. small computer programs using accessible technologies for better distribution and validation. This constitute the \massa\ toolbox, available in public domain in an open Git repository~\cite{gitBook}. The scripts are written in Python and make use of external libraries such as Numpy and Scikits/Audiolab, that performs calls to Fortran routines for better computational efficiency. Part of the scripts have been transcribed to JavaScript and native Python with readiness, what favors their use in Web browsers such as Firefox and Chromium~\cite{numpy, audiolab, tutpython, python}. Furthermore, these are all open technologies, that is, published using licenses that allows copy, distribution and use of any part for research and derivatives. This way, the work presented here is available and eases co-authorship processes~\cite{Raymond,Lessig}.

@@

The remaining parts of this work are organized as follows: section~\ref{sec:dicNote} exposes the presence of sinusoid in discrete-time audio and characterizes a basic musical note; the definition and expansion of the musical note as the essential unit of music is further approached in sections~\ref{sec:discNote} and~\ref{sec:internalVar}, respectively; section~\ref{sec:notesMusic} tackles the organization of musical notes in a higher level~\cite{Wisnick,Webern,Lerdhal,Cook,Lacerda}. Musical theory embody diverse topics as psycho-acoustics, cultural manifestations and formalisms. The section~\ref{sec:results} points out these topics as needed and designate external complements~\cite{Zamacois,Schoenberg,microsound}.

The next section is a minimum text in which music elements are presented side-by-side with the discrete-time samples they result. In order to account for validation and sharing, implementations on computer code of these relations and little musical pieces resulting from them are gathered in the \massa\ toolbox, online available.

>>>>

This article aims at representing musical structures and artifices by their discrete-time sonic characteristics. Results include mathematical relations, usually in terms of samples, and their direct computer program implementations. Despite the general interests involved, there are few books and computer implementations that tackle the subject. These mainly focus on computer implementations and ways to mimic traditional instruments, with scattered mathematical formalisms. A compilation of the works and their contributions is in the bibliography~\cite{dissertacao}. To the best of the author's knowledge, there is a lack of articles on the topic. Moreover, although current computer implementations use the analytical descriptions presented in this study in a implicit manner, it seems that there has been no concise and mathematical description of the processes implemented.

In order to address this concise description of musical elements and structures, in terms of the digitilized sound, the objectives of this paper are:

@@

\item Introduce a framework of sound synthesis with control at sample level and with potential uses in psycho-acoustical experiments and high-fidelity synthesis.

\item Provide the accessibility of the developed framework. The analytic relations presented in this article are implemented as public domain scripts, i.e. small computer programs using accessible technologies for better distribution and validation. This constitute the \massa\ toolbox, available in an open Git repository~\cite{gitBook}. These scripts are written in Python and make use of external libraries, such as Numpy that performs calls to Fortran routines for better computational efficiency. Part of the scripts have been transcribed to JavaScript and native Python with readiness, what favors their use in Web browsers such as Firefox and Chromium~\cite{numpy, audiolab, tutpython, python}. Furthermore, these are all open technologies, that is, published using licenses that allow copy, distribution and use of any part for research and derivatives. This way, the work presented here is embedded in recommended practices for availability and eases co-authorship processes~\cite{Raymond,Lessig}.

\item To provide a didactic presentation of the content to favor its apprehension and usage. It is worthwhile to mention that this subject comprises diverse topics on signal processing, music, psycho-acoustics and programming.

@@

The remaining parts of this work are organized as follows: section~\ref{sec:discNote} characterizes the basic musical note; internals of the musical note is further developed in section~\ref{sec:internalVar}; section~\ref{sec:notesMusic} tackles the organization of musical notes in higher levels of musical structure~\cite{Wisnick,Webern,Lerdhal,Cook,Lacerda}. As these descriptions embody topics such as psycho-acoustics, cultural traditions, formalisms and protocols, the text points to external complements as needed~\cite{Zamacois,Schoenberg,microsound}.

The next section is a minimum text in which musical elements are presented side-by-side with the discrete-time samples they result. In order to account for validation and sharing, implementations on computer code of each one of these relations are gathered in the \massa\ toolbox together with little musical montage resulting from them.
